The role of advection versus diffusion in intraventricular transport

### 2025-03-07 paper/main.tex
update template geometry

%\pagestyle{fancy}
%\rhead{}%\includegraphics[width=2.5cm]{vch-logo.png}}
\pagestyle{plain} % MER

@@

% Affiliations: Please provide adacemic titles (Prof. or Dr.) for all authors where applicable,
% and include an institutional email address for all corresponding authors
\begin{affiliations}
\noindent H.~Herlyng, Dr.~A.~J.~Ellingsrud, Dr.~M.~Kuchta, Dr.~M.~E.~Rognes\\
Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Kristian Augusts gate 23, 0164 Oslo, Norway\\
Email Address: hherlyng@simula.no \\ 
Dr.~M.~E.~Rognes \\
K.~G.~Jebsen Centre for Brain Fluid Research, Oslo, Norway \\
Dr.~I.~Jeong, Dr.~N.~Jurisch-Yaksi\\
Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Erling Skjalgsons gate 1, 7491 Trondheim, Norway \\
Email Address: nathalie.jurisch-yaksi@ntnu.no

\end{affiliations}

@@

@@

Because the elements of $\VV_h$ define the normal components of $\uu_h$,
continuity in tangential components of $\uu_h$ on interior facets of
the mesh is ensured through stabilization
of the discrete weak form~\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}.
We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}.
First, we define the average and jump operators, respectively, as
\begin{equation*}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2}
    \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
\end{equation*}
where $\nn$ is the normal vector on a facet separating two cells,
where the plus and minus signs denote variables in each of the two cells.
The normal vector is chosen such that it points towards the exterior of
its corresponding cell. On the boundary, the operators are defined as
$\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. We now define
and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &=
    \sum_{f\in \mathcal{F}}\int_f{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}
    \cdot\jump{P_{\nn}(\vv_h)}}\\
    &-\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}
    -\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}
\end{align*}
to the discrete weak formulation~\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC},
where $\mathcal{F}$ is the set of all interior facets of the computational mesh.
The first term of $\mathcal{A}_{\mathrm{stab}}$ penalizes jumps
in the tangential components of the velocity
across interior facets, where the penalty parameter $\gamma>0$ must be chosen large enough
to achieve stability. For convergence the penalization term
scales with the cell diameter of a mesh cell $h$,
defined as the maximal distance between two points in a cell. The other two terms
of $\mathcal{A}_{\mathrm{stab}}$ are added to ensure coercivity and symmetry of
the bilinear form, such that the problem is well-defined.

@@

\subsection{Discretization of the advection-diffusion equation}
We outline the discretization of the advection-diffusion weak formulation
for the baseline model, defined
by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.
The weak formulation is discretized with a symmetric interior-penalty method using
discontinuous Galerkin finite elements~\cite{Arnold1982AnElements}.
Let the subscript $h$ denote discrete variables. Exchanging the
variables in~\Cref{eq:adv_diff_weak_formB} with their discrete
counterparts yields the discrete weak form: find $c_h\in W_h$ such that
\begin{equation*}
    a(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h), \quad\forall\phi_h\in W_h.
\end{equation*}
The space $W_h$ is constructed with finite elements with
discontinuous Lagrange polynomial basis functions:
\begin{equation}
    W_h = \Big\{w_h : w_h\in L^2;  w_h\vert_K\in P_k; \forall K\in\Omega_h\Big\}.
\end{equation}
These elements allow for discontinuities in the discrete solution across element edges.
To ensure a stable discretization with these elements, we add the stabilization
\begin{equation}
    \begin{aligned}
        a_{\mathrm{stab}}(c_h, \phi_h)
        = &\sum_{f\in \mathcal{F}}\int_f{D\frac{\alpha}{\avg{h}}\jump{c_h}\cdot\jump{\phi_h}}
        +\sum_{f\in \mathcal{F}}\int_f{\jump{c_h \hat{u}_h}\cdot\jump{\phi_h}}  \\
         -&\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla\phi_h}\cdot\jump{c_h}}
         -\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla c_h}\cdot\jump{\phi_h}},
    \end{aligned}\label{eq:SIP_stabilization}
\end{equation}
to the discrete weak formulation, where
\begin{equation*}
    \hat{u}_h = \frac{1}{2}\left(\uu_h\cdot\nn + \vert\uu_h\cdot\nn\vert\right)
\end{equation*}
is used to upwind the velocity $\uu_h$ with respect to a
facet. Similar to the stabilization employed for the Stokes equations, the first two terms
of \Cref{eq:SIP_stabilization} penalize jumps in the concentration flux across interior facets.
The stabilization parameter $\alpha>0$ must be chosen
large enough to achieve stability. The two last terms in $a_{\mathrm{stab}}$ are
added to preserve coercivity and symmetry of the bilinear form. The average and jump operators used
in~\Cref{eq:SIP_stabilization} are defined, respectively, as
\begin{equation*}
    \avg{c} = \frac{c^+ + c^-}{2} \qquad\mathrm{and}\qquad \jump{c} = c^+\nn^+ + c^-\nn^-.
\end{equation*}
The final discrete weak formulation of the advection-diffusion equation is:
find $c_h\in W_h$, such that
\begin{equation}
    a(c_h, \phi_h) + a_{\mathrm{stab}}(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h),
    \quad\forall\phi_h\in W_h.
    \label{eq:discrete_weak_form_adv_diff}
\end{equation}
In this work, we solved~\Cref{eq:discrete_weak_form_adv_diff} numerically
with the penalty parameter value $\alpha=25$.

>>>>

\pagestyle{plain}

\begingroup
\parindent0pt

@@

% Affiliations: Please provide adacemic titles (Prof. or Dr.) for all authors where applicable,
% and include an institutional email address for all corresponding authors
\begin{affiliations}
H.~Herlyng, Dr.~A.~J.~Ellingsrud, Dr.~M.~Kuchta, Dr.~M.~E.~Rognes\\
Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Kristian Augusts gate 23, 0164 Oslo, Norway\\
Email Address: hherlyng@simula.no

\medskip
Dr.~M.~E.~Rognes \\
K.~G.~Jebsen Centre for Brain Fluid Research, Oslo, Norway

\medskip
Dr.~I.~Jeong, Dr.~N.~Jurisch-Yaksi\\
Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Erling Skjalgsons gate 1, 7491 Trondheim, Norway \\
Email Address: nathalie.jurisch-yaksi@ntnu.no
\end{affiliations}

\medskip

@@

\endgroup % End 0parindent group

@@

Because the elements of $\VV_h$ define the normal components of $\uu_h$,
continuity in tangential components of $\uu_h$ on interior facets of
the mesh is ensured through stabilization
of the discrete weak form~\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC}.
We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}.
First, we define the average and jump operators, respectively, as
\begin{equation*}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2}
    \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
\end{equation*}
where $\nn$ is the normal vector on a facet separating two cells,
where the plus and minus signs denote variables in each of the two cells.
The normal vector is chosen such that it points towards the exterior of
its corresponding cell. On the boundary, the operators are defined as
$\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. We now define
and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &=
    \sum_{f\in \mathcal{F}}\int_f{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}
    \cdot\jump{P_{\nn}(\vv_h)}}\,\mathrm ds\\
    &-\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}\,\mathrm ds
    -\sum_{f\in \mathcal{F}}\int_f{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}\,\mathrm ds
\end{align*}
to the discrete weak formulation~\Cref{eq:no_stabilization_abstract_discrete_weak_form_modelC},
where $\mathcal{F}$ is the set of all interior facets of the computational mesh.
The first term of $\mathcal{A}_{\mathrm{stab}}$ penalizes jumps
in the tangential components of the velocity
across interior facets, where the penalty parameter $\gamma>0$ must be chosen large enough
to achieve stability. For convergence the penalization term
scales with the cell diameter of a mesh cell $h$,
defined as the maximal distance between two points in a cell. The other two terms
of $\mathcal{A}_{\mathrm{stab}}$ are added to ensure coercivity and symmetry of
the bilinear form, such that the problem is well-defined.

@@

\subsection{Discretization of the advection-diffusion equation}
We outline the discretization of the advection-diffusion weak formulation
for the baseline model, defined
by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.
The weak formulation is discretized with a symmetric interior-penalty method using
discontinuous Galerkin finite elements~\cite{Arnold1982AnElements}.
Let the subscript $h$ denote discrete variables. Exchanging the
variables in~\Cref{eq:adv_diff_weak_formB} with their discrete
counterparts yields the discrete weak form: find $c_h\in W_h$ such that
\begin{equation*}
    a(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h), \quad\forall\phi_h\in W_h.
\end{equation*}
The space $W_h$ is constructed with finite elements with
discontinuous Lagrange polynomial basis functions:
\begin{equation}
    W_h = \Big\{w_h : w_h\in L^2;  w_h\vert_K\in P_k; \forall K\in\Omega_h\Big\}.
\end{equation}
These elements allow for discontinuities in the discrete solution across element edges.
To ensure a stable discretization with these elements, we add the stabilization
\begin{equation}
    \begin{aligned}
        a_{\mathrm{stab}}(c_h, \phi_h)
        = &\sum_{f\in \mathcal{F}}\int_f{D\frac{\alpha}{\avg{h}}\jump{c_h}\cdot\jump{\phi_h}}\,\mathrm ds
        +\sum_{f\in \mathcal{F}}\int_f{\jump{c_h \hat{u}_h}\cdot\jump{\phi_h}}\,\mathrm ds  \\
         -&\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla\phi_h}\cdot\jump{c_h}}\,\mathrm ds
         -\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla c_h}\cdot\jump{\phi_h}}\,\mathrm ds,
    \end{aligned}\label{eq:SIP_stabilization}
\end{equation}
to the discrete weak formulation, where
\begin{equation*}
    \hat{u}_h = \frac{1}{2}\left(\uu_h\cdot\nn + \vert\uu_h\cdot\nn\vert\right)
\end{equation*}
is used to upwind the velocity $\uu_h$ with respect to a
facet. Similar to the stabilization employed for the Stokes equations, the first two terms
of \Cref{eq:SIP_stabilization} penalize jumps in the concentration flux across interior facets.
The stabilization parameter $\alpha>0$ must be chosen
large enough to achieve stability. The two last terms in $a_{\mathrm{stab}}$ are
added to preserve coercivity and symmetry of the bilinear form. The average and jump operators used
in~\Cref{eq:SIP_stabilization} are defined, respectively, as
\begin{equation*}
    \avg{c} = \frac{c^+ + c^-}{2} \qquad\mathrm{and}\qquad \jump{c} = c^+\nn^+ + c^-\nn^-.
\end{equation*}
The final discrete weak formulation of the advection-diffusion equation is:
find $c_h\in W_h$, such that
\begin{equation}
    a(c_h, \phi_h) + a_{\mathrm{stab}}(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h),
    \quad\forall\phi_h\in W_h.
    \label{eq:discrete_weak_form_adv_diff}
\end{equation}
In this work, we solved~\Cref{eq:discrete_weak_form_adv_diff} numerically
with the penalty parameter value $\alpha=25$.
